A geo-referenced micro-data set of real estate listings for Spain’s three largest cities

### 2023-05-19 EPB/main_EPB.tex
Resubmission to EPB

\begin{abstract}
This article presents an open data product with large geo-referenced
micro-data sets of 2018 real estate listings in Spain. These data were
originally published on the idealista.com real estate website. The
observations were obtained for the three largest cities in Spain: Madrid
(n = 94,815 observations), Barcelona (n = 61,486 observations), and
Valencia (n = 33,622 observations). The data sets include the
coordinates of properties (latitude and longitude), asking prices of
each listed dwelling, and several variables of indoor characteristics.
The listings were enriched with official information from the Spanish
cadastre (e.g., building material quality) plus other relevant
geographical features, such as distance to urban points of interest.
Along with the real estate listings, the data product also includes
neighborhood boundaries for each city. The data product is offered as a
fully documented R package and is available for scientific and
educational purposes, particularly for geo-spatial studies
\end{abstract}

@@

Interest in the characteristics of the housing market and housing prices
has been a growing area of research in recent decades, generating a vast
amount of theoretical and empirical literature. Including the spatial
component to analyze the real estate market and incorporating geographic
variables has significantly improved the understanding of this market:
to this end, it is essential to have information/data at the point
level. Therefore, it is becoming common for spatial analysis of urban
environments to be use micro-data, geo-referenced as points
\citep{lopez2015}. In some cases, researchers have had to resort to
webscraping processes to obtain data for research
\citep[e.g.,][]{gupta2022take, arbia2020spatial, Li2019, lopez2015}.
Alas, webscraping is a chancy process prone to download errors, missing
data, duplicated records, etc. Furthermore, researchers do not always
share their webscrapped datasets, which limits reproducibility of their
research. As we witness a growing interest in openness and
reproducibility in geographical data science
\citep{arribasl2021editorial, paez2021open, brunsdon2021opening}, it
becomes increasingly urgent to have open data products to support
research \citep{arribas2021}.

@@

To contribute to the growing inventory of international micro-data sets
of real estate markets, this paper introduces an open micro-data set of
geo-referenced dwelling listings. The data have been provided by the
Idealista
company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal}
and contain information about 189,923 dwellings located in Spain's three
largest cities. To date, this data product is one of the largest open
geo-referenced micro-data set of the housing market in the world. The
most similar in terms of geographical disaggregation and sample size is
the dataset of \citet{song2021hedonic} which includes transactions for
four cities in South Korea, namely Busan (\(n=61,152\)), Daegu
(\(n=32,363\)), Daejeon (\(n=21,114\)), and Gwangju (\(n=25,984\)). It
is certainly the largest in Spain.

The data set has been supplied directly by Idealista, and therefore is
clean and free of download errors. However, to comply with data
legislation, to share the data we have slightly modified the prices by
applying a small amount of random noise which will not bias the main
results derived from its usage.

This micro-data set can be used to benchmark new analytical methods in a
reproducible fashion. Applied and theoretical researchers on real estate
mass appraisal and valuation methods might use this dataset to
canonically compare the performance of their proposed comparable and
hedonic models, among others. The data can also be used to study the
segmentation of housing submarkets and related topics such as the impact
of suburban areas on house prices. The listings have been enriched with
official information from the Spanish cadastre along with other relevant
geographical features, such as distance to urban points of interest. In
any case, the data might be easily extended by spatially joining other
datasets that contain information at administrative, census tract, or
postal code levels.

The data set is distributed as an \texttt{R} package, named
\{idealista18\}, which can be accessed from a public Github
repository\footnote{\url{https://github.com/paezha/idealista18}}. The
open data product is made available under the
\href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
For transparency, we also share the randomization process applied to the
original data in the aforementioned Github repository.

@@

\begin{table}[!ht]
    \centering
      \caption{Total properties and transactionst three Spanish cities. Year 2018} 
  \label{tab:transactions} 
    \begin{tabular}{lccccc}
        City & Total properties (P) & Total transactions (T) & Listings (L) & L/P & T/L \\
        \hline
        Barcelona &  789,740  &  56,012  &  61,329  & 7.8\% & 81.3\%  \\ 
        Madrid &  1,545,397  &  76,603  &  94,802  & 6.1\%  & 80.8\% \\ 
        Valencia &  416,004  &  30,615  &  33,593  & 8.1\% & 91.1\% \\ 
        \hline
        Total &  2,751,141  &  163,230  &  189,724  & 6.9\% & 86.0\% \\ 
        \hline
        \multicolumn{6}{l}{{\footnotesize \textbf{Sources}}}\\
        \multicolumn{6}{l}{{\footnotesize Total properties (P): Ministerio Español de Hacienda y Función Pública}}\\
        \multicolumn{6}{l}{{\footnotesize Total transactions (T): Instituto Nacional de Estadística}}\\
    \end{tabular}
\end{table}

The following subsections describe each object. A full description of
the data is available in the help section of the package. We have tried
to the best of our ability to comply with FAIR principles regarding
research data \citep{Wilkinson2016fair}: upon publication, the dataset
has a persistent digital object identifier; publication as a data
article makes the data findable; the data and metadata are packaged
together and protocols for help files in the R ecosystem mean that
documentation is easily searchable; distribution as an R package means
that only open software is needed to access the data; and a public
repository documents all data processes followed to generate the
distributed open data product.

@@

The dwelling listing of each city includes a set of characteristics for
each dwelling published on the idealista real estate website as an ad.
The dwelling listing has been included in the \{idealista18\} package as
a simple features (sf) object \citep{Pebesma} with point geometry in
latitude and longitude. The name of each sf object with the list of
dwellings is the name of the city followed by '\_Sale' (e.g.,
Madrid\_Sale). Each data object includes a total of 42 variables and the
complete set of listings corresponding to the four quarters of 2018.
Table \ref{tab:number-ads} shows the number of dwelling listing ads
included in the data set for each city and quarter. The record counts
for each city are: 94,815 listings for Madrid, 61,486 for Barcelona, and
33,622 for Valencia. Note that the same dwelling may be found in more
than one period when a property listed for sale in one quarter was sold
in a subsequent quarter. The variable ASSETID, included in the sf
objects, is the unique identifier of the dwelling.

@@

Each record of the dwelling listing contains a set of indoor
characteristics supplied by the advertisers on the Idealista website
(e.g., price, surface area, number of bedrooms, basic features, etc.),
including the location of the dwelling. The exact location has been
masked, as described in Section
\protect\hyperlink{anonymizing}{Anonymizing the data set}). Table
\ref{tab:variables} lists the main indoor variables with a short
description and the mean value of each variable. The dwelling listings
were enriched with a number of additional attributes from the Spanish
cadastre \citep{Catastro}. The cadastral information is described in
Table \ref{tab:variables}, with the prefix CAD in the variable name. The
cadastral features were assigned by applying the features of the nearest
parcel to the coordinates. The year the dwelling was built
(CONSTRUCTIONYEAR) given by the advertiser was revised since the year of
construction is entered on the website by users, and it is therefore
subject to errors and incomplete data (40\% missing data). To resolve
this issue, an alternative variable (CADCONSTRUCTIONYEAR) was included,
assigning the cadastral construction year from the nearest cadastral
parcel whenever the value was outstanding (date was after publication
date or year of construction was before 1500) or when the value supplied
by the advertiser was missing.

@@

\begin{table}[ht]
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}{>{\raggedright\arraybackslash}p{13em}>{\raggedright\arraybackslash}p{14em}ccc}
  \hline
Variable & Sort Description & Barcelona & Madrid & Valencia \\ 
  \hline
PRICE & Asking price & 395770.58 & 396110.11 & 199678.31 \\ 
  UNITPRICE & Asking price per m\verb|^|2 (euros) & 4044.86 & 3661.05 & 1714.54 \\ 
  CONSTRUCTEDAREA & Surface (m\verb|^|2) & 95.46 & 101.40 & 108.95 \\ 
  ROOMNUMBER & Number of bedrooms & 2.86 & 2.58 & 3.07 \\ 
  BATHNUMBER & Number of bathrooms & 1.52 & 1.59 & 1.59 \\ 
  CONSTRUCTIONYEAR & Construction year (advertiser) & 1952.58 & 1964.69 & 1969.43 \\ 
  CADCONSTRUCTIONYEAR & Construction year (cadastre) & 1952.19 & 1965.70 & 1970.55 \\ 
  CADMAXBUILDINGFLOOR & Max build floor & 6.85 & 6.38 & 7.04 \\ 
  CADDWELLINGCOUNT & Dwelling count in the building & 28.56 & 39.19 & 36.83 \\ 
  CADASTRALQUALITYID & Cadastral quality. 0 Best-10 Worst & 4.31 & 4.85 & 5.34 \\ 
  DISTANCE\_TO\_CITY\_CENTER & Distance to city center & 2.80 & 4.49 & 2.09 \\ 
  DISTANCE\_TO\_METRO & Distance to subway station & 0.27 & 0.48 & 0.64 \\ 
  DISTANCE\_TO\_(MAINSTREET) & Distance to major street & 1.77 & 2.68 & 2.07 \\ 
   \hline
\end{tabular}
\caption{List, sort description, and mean of the main quantitative variables included in the dwelling listing for the three Spanish cities. See the help section in the \textbf{idealista18} R package for details and formal definitions. Some variables have been excluded from this table to save space. Check the full list in the \textbf{idealista18} package.\label{tab:variables}} 
\end{table}

@@

The second data object included in \{idealista18\} is the neighborhoods
in the three cities as polygons. There is an sf object for each city
with the name of the city and the suffix '\_Polygons'. Figure
\ref{fig:all-polygons} shows the quantile maps of the number of
dwellings in the listing for the different neighborhoods in the three
cities. The neighborhoods are based on the official boundaries but
slightly changed by
Idealista\footnote{There are two criteria used to make this division. If an area is small enough and similar enough to another, the two areas are merged, and, if the official area is not homogeneous, it is divided into a series of new polygons.}.
In practical terms, we can assume they are the same since the website
combines areas when there are few ads for that area. In the case of
Madrid, they combined four areas into two.

\begin{figure}
\centering
\includegraphics{main_EPB_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:all-polygons}Number of listings in each
neighborhood. Boundary for Barcelona (Top), Madrid (Center), and
Valencia (Bottom).}
\end{figure}

@@

The total number of dwellings is available from the Spanish cadastre
aggregated by district (districts are groups of neighborhoods). Figure
\ref{fig:districts} shows the percentage of listed dwellings relative to
the total number of dwelling by district in each of the three cities.
This gives a sense of how active residential real estate markets were in
different parts of each city in 2018.

\begin{figure}
\centering
\includegraphics{main_EPB_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{\label{fig:districts}Percentage of listings relative to total
number of dwellings by district. Boundary for Barcelona (Top), Madrid
(Center), and Valencia (Bottom).}
\end{figure}

@@

\hypertarget{anonymizing}{%
\section{Anonymizing the data set}\label{anonymizing}}

To comply with Spanish regulations, two variables were slightly modified
to provide anonymity. A masking process was applied to asking prices and
location (coordinates).

@@

\begin{algorithm}[!ht]
 \KwData{all idealista listings}
 \KwResult{all idealista listings with masked coordinates}
 initialization\;
 \For{each listing L}{
  take geographical location of L as $(X,Y)$
  \Repeat{this stop condition}{
    take a random angle $\alpha$ from 0 to 360 degrees
    take a distance $R$ as a random value from 30 to 60 meters
    determine a new point $(X',Y')$ calculated as a point located $R$ with the angle $\alpha$
  }
  set $(X',Y')$ as the new location for the listing L
 }
 \caption{Coordinate displacement process for anonymisation purposes}
 \label{algo:coordinates-displacement}
\end{algorithm}

@@

@@

\caption{\label{fig:Anonymizing}(Left) Masking coordinates. Spatial range. (Right) Coordinate displacement in meters (Valencia)}\label{fig:unnamed-chunk-3}
\end{figure}

@@

This paper describes a data product of a geo-referenced micro-data set
of Spain's three largest cities. This is an excellent data product to
help understand the complex mechanisms related to the housing market and
housing prices. Researchers can apply hedonic models with spatial
effects, identifying housing submarkets or machine learning techniques
\citep[e.g.,][]{rey2023using}. The data product can also be used for
educational proposes and teaching activities.

@@

Author One and author Two are employed by Idealista. They have been
granted permission to share the data presented in this article. None of
the authors have financial interests or personal relationships which
have, or could be perceived to have, influenced the work reported in
this article.

@@

The authors wish to thank Alessandro Galesi for their support in the
paper revision and Juan Ramón Selva for collecting and cleaning the
spatial data. This work has been partially funded by the Spanish
Ministry of Economy and Competitiveness Grants PID2019-107800GB-100

>>>>

\begin{abstract}
This article presents an open data product with large geo-referenced
micro-data sets of 2018 real estate listings in Spain. These data were
originally published on the idealista.com real estate website. The
observations were obtained for the three largest cities in Spain: Madrid
(n = 94,815 observations), Barcelona (n = 61,486 observations), and
Valencia (n = 33,622 observations). The data sets include the
coordinates of properties (latitude and longitude), asking prices of
each listed dwelling, and several variables of indoor characteristics.
The listings were enriched with official information from the Spanish
cadastre (e.g., building material quality) plus other relevant
geographical features, such as distance to urban points of interest.
Along with the real estate listings, the data product also includes
neighborhood boundaries for each city. The data product is offered as a
fully documented \texttt{R} package and is available for scientific and
educational purposes, particularly for geo-spatial studies
\end{abstract}

@@

Interest in the characteristics of the housing market and housing prices
has been a growing area of research in recent decades, generating a vast
amount of theoretical and empirical literature. Including the spatial
component to analyze the real estate market and incorporating geographic
variables has significantly improved the understanding of this market:
to this end, it is essential to have information/data at the point
level. Therefore, it is becoming common for spatial analysis of urban
environments to be use micro-data, geo-referenced as points
\citep{lopez2015}. In some cases, researchers have had to resort to
webscraping processes to obtain data for research
\citep[e.g.,][]{Li2019, lopez2015}. Alas, webscraping is a chancy
process prone to download errors, missing data, duplicated records, etc.
Furthermore, researchers do not always share their webscrapped datasets,
which limits reproducibility of their research. As we witness a growing
interest in openness and reproducibility in geographical data science
\citep{arribasl2021editorial, paez2021open, brunsdon2021opening}, it
becomes increasingly urgent to have open data products to support
research \citep{arribas2021}.

@@

To contribute to the growing inventory of international micro-data sets
of real estate markets, this paper introduces an open data product with
geo-referenced dwelling listings, called \{idealista18\}. The data have
been provided by the Idealista
company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal}
and contain information about 189,923 dwellings located in Spain's three
largest cities. To date, this data product is one of the largest open
geo-referenced micro-data set of the housing market in the world. The
most similar in terms of geographical disaggregation and sample size is
the dataset of \citet{song2021hedonic} which includes transactions for
four cities in South Korea, namely Busan (\(n=61,152\)), Daegu
(\(n=32,363\)), Daejeon (\(n=21,114\)), and Gwangju (\(n=25,984\)).
\{idealista18\} is certainly the largest open data product of its kind
in Spain.

The data set has been supplied directly by Idealista, and therefore is
clean and free from download errors. However, to comply with data
legislation, we have masked the prices by applying a small amount of
random noise that will not bias the main results derived from its usage.

This micro-data set can be used to benchmark new methods in a
reproducible fashion \citep[e.g.,][]{rey2023using}. Applied and
theoretical researchers on real estate mass appraisal and valuation
methods might use this dataset to canonically compare the performance of
their proposed comparable and hedonic models, among others. The data can
also be used to study the segmentation of housing submarkets and related
topics such as the impact of suburban areas on house prices. The
listings have been enriched with official information from the Spanish
cadastre along with other relevant geographical features, such as
distance to urban points of interest. In any case, the data might be
easily extended by spatially joining other datasets that contain
information at administrative, census tract, or postal code levels.

The data set is distributed as an \texttt{R} package and it can be
accessed from a public Github
repository\footnote{\url{https://github.com/paezha/idealista18}}. The
open data product is made available under the
\href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
For transparency, we also share the masking process applied to the
original data in the aforementioned Github repository.

@@

\begin{table}[!ht]
\centering
\caption{Total properties and transactionst three Spanish cities. Year 2018} 
\label{tab:transactions} 
\begin{tabular}{lccccc}
City & Total properties (TP) & Total transactions (TT) & Listings (L) & L/TP & TT/L \\
\hline
Barcelona &  789,740  &  56,012  &  61,329  & 7.8\% & 81.3\%  \\ 
Madrid &  1,545,397  &  76,603  &  94,802  & 6.1\%  & 80.8\% \\ 
Valencia &  416,004  &  30,615  &  33,593  & 8.1\% & 91.1\% \\ 
\hline
Total &  2,751,141  &  163,230  &  189,724  & 6.9\% & 86.0\% \\ 
\hline
\multicolumn{6}{l}{{\footnotesize \textbf{Sources}}}\\
\multicolumn{6}{l}{{\footnotesize Total properties (P): Ministerio Español de Hacienda y Función Pública}}\\
\multicolumn{6}{l}{{\footnotesize Total transactions (T): Instituto Nacional de Estadística}}\\
\end{tabular}
\end{table}

The following subsections describe the data objects. A full description
of the data is also available in the help section of the package. We
have tried to the best of our ability to comply with FAIR principles
regarding research data \citep{Wilkinson2016fair}: upon publication, the
dataset has a persistent digital object identifier; publication as a
data article makes the data findable; the data and metadata are packaged
together and protocols for help files in the \texttt{R} ecosystem mean
that documentation is easily searchable; distribution as an \texttt{R}
package means that only open software is needed to access the data; and
a public repository documents all data processes followed to generate
the distributed open data product.

@@

The listing for each city includes a set of characteristics for each
dwelling published on the idealista real estate website. The listing
included in the \{idealista18\} package are simple features (sf) objects
\citep{Pebesma} with point geometry in latitude and longitude. The name
of each sf object with the list of dwellings is the name of the city
followed by '\_Sale' (e.g., \texttt{Madrid\_Sale}). Each data object
includes a total of 42 variables and the complete set of listings
corresponding to the four quarters of 2018 (Q1 through Q4). Table
\ref{tab:number-ads} shows the number of dwelling listing ads included
in the data set for each city and quarter. The record counts for each
city are: 94,815 listings for Madrid, 61,486 for Barcelona, and 33,622
for Valencia. Note that the same dwelling may be found in more than one
period when a property listed for sale in one quarter was sold in a
subsequent quarter. The variable ASSETID, included in the sf objects, is
the unique identifier of the dwelling.

@@

Each record of the dwelling listing contains a set of indoor
characteristics supplied by the advertisers on the Idealista website
(e.g., price, surface area, number of bedrooms, basic features, etc.),
including an approximated location of the dwelling (the exact location
has been masked, as described in Section
\protect\hyperlink{anonymizing}{Masking the prices}). Table
\ref{tab:variables} lists the main indoor variables with a short
description and the mean value of each variable. The dwelling listings
were enriched with a number of additional attributes from the Spanish
cadastre \citep{Catastro}. The cadastral information is described in
Table \ref{tab:variables}, with the prefix CAD in the variable name. The
cadastral features were assigned by applying the features of the nearest
parcel to the coordinates. The year the dwelling was built
(CONSTRUCTIONYEAR) given by the advertiser was revised since the year of
construction is entered on the website by users, and it is therefore
subject to errors and incomplete data (40\% missing data). To resolve
this issue, an alternative variable (CADCONSTRUCTIONYEAR) was included,
assigning the cadastral construction year from the nearest cadastral
parcel whenever the value was outstanding (date was after publication
date or year of construction was before 1500) or when the value supplied
by the advertiser was missing.

@@

\begin{table}[ht]
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}{>{\raggedright\arraybackslash}p{13em}>{\raggedright\arraybackslash}p{14em}ccc}
  \hline
Variable & Sort Description & Barcelona & Madrid & Valencia \\ 
  \hline
PRICE & Asking price & 395770.58 & 396110.11 & 199678.31 \\ 
  UNITPRICE & Asking price per m\verb|^|2 (euros) & 4044.86 & 3661.05 & 1714.54 \\ 
  CONSTRUCTEDAREA & Surface (m\verb|^|2) & 95.46 & 101.40 & 108.95 \\ 
  ROOMNUMBER & Number of bedrooms & 2.86 & 2.58 & 3.07 \\ 
  BATHNUMBER & Number of bathrooms & 1.52 & 1.59 & 1.59 \\ 
  CONSTRUCTIONYEAR & Construction year (advertiser) & 1952.58 & 1964.69 & 1969.43 \\ 
  CADCONSTRUCTIONYEAR & Construction year (cadastre) & 1952.19 & 1965.70 & 1970.55 \\ 
  CADMAXBUILDINGFLOOR & Max build floor & 6.85 & 6.38 & 7.04 \\ 
  CADDWELLINGCOUNT & Dwelling count in the building & 28.56 & 39.19 & 36.83 \\ 
  CADASTRALQUALITYID & Cadastral quality. 0 Best-10 Worst & 4.31 & 4.85 & 5.34 \\ 
  DISTANCE\_TO\_CITY\_CENTER & Distance to city center & 2.80 & 4.49 & 2.09 \\ 
  DISTANCE\_TO\_METRO & Distance to subway station & 0.27 & 0.48 & 0.64 \\ 
  DISTANCE\_TO\_(MAINSTREET) & Distance to major street & 1.77 & 2.68 & 2.07 \\ 
   \hline
\end{tabular}
\caption{List, sort description, and mean of the main quantitative variables included in the dwelling listing for the three Spanish cities. See the help section in the \{idealista18\} package for details and formal definitions. Some variables have been excluded from this table to save space. Check the full list in the package.\label{tab:variables}} 
\end{table}

@@

The second data object included in \{idealista18\} is the neighborhoods
in the three cities as polygons. There is an sf object for each city
with the name of the city and the suffix '\_Polygons'. The left column
of Figure \ref{fig:all-polygons} shows the quantile maps of the number
of dwellings in the listing for the different neighborhoods in the three
cities. The neighborhoods are based on the official boundaries but
slightly changed by
Idealista\footnote{There are two criteria used to make this division. If an area is small enough and similar enough to another, the two areas are merged, and, if the official area is not homogeneous, it is divided into a series of new polygons.}.
In practical terms, we can assume they are the same since the website
combines areas when there are few ads for that area. In the case of
Madrid, they combined four areas into two.

@@

The total number of dwellings is available from the Spanish cadastre
aggregated by district (districts are groups of neighborhoods). The
right column in Figure \ref{fig:all-polygons} shows the percentage of
listed dwellings relative to the total number of dwelling by district in
each of the three cities. This gives a sense of how active residential
real estate markets were in different parts of each city in 2018.

\begin{figure}
\centering
\includegraphics{main_EPB_files/figure-latex/figure-listings-1.pdf}
\caption{\label{fig:all-polygons}Listings by neighborhood (left column)
and percentage of listings relative to total dwellings (right column).
Barcelona (Top), Madrid (Center), and Valencia (Bottom).}
\end{figure}

@@

\hypertarget{anonymizing}{%
\section{Masking the prices}\label{anonymizing}}

To comply with Spanish regulations, three variables were slightly
modified to provide a measure of anonymity. A masking process was
applied to asking prices and location (coordinates).

@@

To understand the implications of this masking process, we use some
standard results from algebra of random variables. The masked prices
\(P\) are given by: \[
P = RP\cdot \epsilon
\]

\noindent where \(RP\) are the original (raw) prices, and \(\epsilon\)
is a random variable drawn from the uniform distribution with parameters
\(a=0.975\) and \(b=1.025\): \[
f(\epsilon) = 
\begin{cases}
\frac{1}{b - a} & \text{for } a\le \epsilon \le b\\
0 & \text{otherwise}
\end{cases}
\]

The expectation of \(\epsilon\) given these parameters is: \[
E[\epsilon] = \frac{a+b}{2} = \frac{0.975 + 1.025}{2} = 1
\]

\noindent and the variance of \(\epsilon\) is: \[
\mathrm{V}[\epsilon]=\frac{(b - a)^2}{12} = \frac{1}{4800}
\]

Therefore, the expectation of the masked prices is: \[
\mathrm{E}[P] = \mathrm{E}[RP\cdot \epsilon] = \mathrm{E}[RP]\cdot \mathrm{E}[\epsilon] = \mathrm{E}[RP]
\]

In other words, the masked prices \(P\) are an unbiased version of the
raw prices \(RP\).

Considering that \(RP\) and \(\epsilon\) are independent, the variance
of the masked prices is as follows: \[
\mathrm{V}[P] = \mathrm{V}[RP\cdot \epsilon] = \mathrm{V}[RP]\cdot \mathrm{V}[\epsilon] + \mathrm{V}[RP]\cdot (\mathrm{E}[\epsilon])^2 + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2
\]

Since \(\mathrm{E}[\epsilon] = 1\), we have that: \[
\mathrm{V}[P] = \mathrm{V}[RP]\cdot \mathrm{V}[\epsilon] + \mathrm{V}[RP] + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2 = \mathrm{V}[RP]\cdot (1 + \mathrm{V}[\epsilon]) + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2
\]

Solving for \(\mathrm{V}[RP]\), and replacing the expectation of the raw
prices by its unbiased version (\(\mathrm{E}[P]\)), yields the variance
of the raw prices: \[
\mathrm{V}[RP] = \frac{\mathrm{V}[P] - \frac{1}{4800}\cdot (\mathrm{E}[P])^2}{1 + \frac{1}{4800}}
\]

Table \ref{tab:variance} reports the mean and the standard deviation
(i.e., the square root of the variance) of the prices in the package,
and the standard deviation (again, the square root of the variance) of
the raw prices calculated using the formula above. This is done for each
quarter and for the full year. The last column of the table can be read
as an inflation factor. We see that the variance of the masked prices is
inflated with respect to the variance of the original values by less
than \(1\%\) in all cases examined. Users can use the formula above to
calculate the inflation of the variance if they use subsamples other
than those shown, to assess the potential impacts of the masking (e.g.,
when computing intervals of confidence).

\begin{table}

\caption{\label{tab:variance-inflation-calculations}\label{tab:variance}Inflation of the variance of masked prices with respect to raw prices}
\centering
\fontsize{9}{11}\selectfont
\begin{threeparttable}
\begin{tabular}[t]{llllll}
\toprule
City & Period & mean(P) & sd(P) & sd(RP) & sd(P)/sd(RP)\\
\midrule
 & Q1 & 405,166.8 & 308,623.8 & 308,536.2 & 1.00057\\
\cmidrule{2-6}
 & Q2 & 388,053.5 & 252,520.5 & 252,432.1 & 1.0007\\
\cmidrule{2-6}
 & Q3 & 382,692.5 & 268,880.9 & 268,796.1 & 1.00063\\
\cmidrule{2-6}
 & Q4 & 398,157.8 & 275,459.3 & 275,370.6 & 1.00064\\
\cmidrule{2-6}
\multirow{-5}{*}{\raggedright\arraybackslash BARCELONA} & 2018 & 395,770.6 & 281,554.8 & 281,467.5 & 1.00062\\
\cmidrule{1-6}
 & Q1 & 404,960.8 & 447,935.3 & 447,850.5 & 1.00038\\
\cmidrule{2-6}
 & Q2 & 367,527.5 & 383,093.9 & 383,017.2 & 1.0004\\
\cmidrule{2-6}
 & Q3 & 365,467.2 & 359,118.3 & 359,042.2 & 1.00042\\
\cmidrule{2-6}
 & Q4 & 410,952.7 & 428,852.7 & 428,767 & 1.0004\\
\cmidrule{2-6}
\multirow{-5}{*}{\raggedright\arraybackslash MADRID} & 2018 & 396,110.1 & 417,074.4 & 416,991.8 & 1.0004\\
\cmidrule{1-6}
 & Q1 & 204,836.5 & 187,957.4 & 187,914.6 & 1.00046\\
\cmidrule{2-6}
 & Q2 & 176,661 & 141,002.4 & 140,964.6 & 1.00054\\
\cmidrule{2-6}
 & Q3 & 188,189.4 & 167,146 & 167,106.6 & 1.00047\\
\cmidrule{2-6}
 & Q4 & 208,523.5 & 183,452.7 & 183,409 & 1.00048\\
\cmidrule{2-6}
\multirow{-5}{*}{\raggedright\arraybackslash VALENCIA} & 2018 & 199,678.3 & 177,156 & 177,114.1 & 1.00047\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item \textit{Note: } 
\item P: masked prices; 
\item RP: raw prices; 
\item sd: standard deviation (square root of the variance)
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\begin{algorithm}[!ht]
 \KwData{all idealista listings}
 \KwResult{all idealista listings with masked coordinates}
 initialization\;
 \For{each listing L}{
  take geographical location of L as $(X,Y)$
  \Repeat{this stop condition}{
    take a random angle $\alpha$ from 0 to 360 degrees
    take a distance $R$ as a random value from 30 to 60 meters
    determine a new point $(X',Y')$ calculated as a point located $R$ with the angle $\alpha$
  }
  set $(X',Y')$ as the new location for the listing L
 }
 \caption{Coordinate displacement process for masking purposes}
 \label{algo:coordinates-displacement}
\end{algorithm}

@@

\caption{\label{fig:Anonymizing}(Left) Masking coordinates. Spatial range. (Right) Coordinate displacement in meters (Valencia)}\label{fig:figure-masking-coordinates}
\end{figure}

@@

This paper describes a data product of a geo-referenced micro-data set
of Spain's three largest cities. This data product can be of value to
support research into the mechanisms of housing market and housing
prices. Researchers can apply hedonic models with spatial effects,
identify housing submarkets, or experiment with machine learning
techniques. The data product can also be used for educational proposes
and teaching activities. To the best of our knowledge, this is the
largest, publicly available data set of its type that is also analysis
ready and fully documented.

@@

Authors One and Two are employed by Idealista. They have been granted
permission to share the data presented in this article. None of the
authors have financial interests or personal relationships which have,
or could be perceived to have, influenced the work reported in this
article.

@@

The authors wish to thank Alessandro Galesi for their support in the
paper revision and Juan Ramón Selva for collecting and cleaning the
spatial data. This work has been partially funded by the Spanish
Ministry of Economy and Competitiveness Grants PID2019-107800GB-100, but
it was not funded by any of Canada's research councils.

### 2023-05-19 EPB/main_EPB_v0.Rmd
Resubmission to EPB

---
title: A geo-referenced micro-data set of real estate listings for Spain’s three largest cities
runninghead: Rey-Blanco \emph{et al}.
author:
- name: D. Rey-Blanco
  num: 1
- name: P. Arbues
  num: 1
- name: F. López
  num: 2
- name: A. Páez*
  num: 3
address:
- num: 1
  org: Idealista, Plaza de las Cortes 5, 28014 Madrid, Spain
- num: 2
  org: Facultad de CC de la Empresa, C/ Real, 3. 30201 Cartagena, Murcia, Spain
- num: 3
  org: School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada
corrauth: "Antonio Páez, School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada."
email: paezha@mcmaster.ca
abstract: "This article shares an open data product with big geo-referenced micro-data sets of 2018 real estate listings in Spain. These data were originally published on the idealista.com real estate website. The observations were obtained for the three largest cities in Spain: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations), and Valencia (n = 33,622 observations). The data sets include the coordinates of properties (latitude and longitude), asking prices of each listed dwelling, and several variables of indoor characteristics. The listings were enriched with official information from the Spanish cadastre (e.g., building material quality) plus other relevant geographical features, such as distance to urban points of interest. Along with the real estate listings, the data product also includes neighborhood boundaries for each city. The data product is offered as a fully documented R package and is available for scientific and educational purposes, particularly for geo-spatial studies"
keywords: Housing market; idealista.com; geo-referenced data; point-level data; open data; Spain
classoption:
  - Royal
  - times
bibliography: bibEPB.bib
bibliographystyle: sageh

@@

The data set is distributed as an R package, name {idealista18}, which can be accessed from a public Github repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}. The open data product {idealista18} is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}. For transparency, we also share the randomisation process applied to the original data in the aforementioned Github repository.

>>>>

---
title: A geo-referenced micro-data set of real estate listings for Spain’s three largest cities
runninghead: author \emph{et al}.
author:
- name: 
abstract: "This article shares an open data product with big geo-referenced micro-data sets of 2018 real estate listings in Spain. These data were originally published on the idealista.com real estate website. The observations were obtained for the three largest cities in Spain: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations), and Valencia (n = 33,622 observations). The data sets include the coordinates of properties (latitude and longitude), asking prices of each listed dwelling, and several variables of indoor characteristics. The listings were enriched with official information from the Spanish cadastre (e.g., building material quality) plus other relevant geographical features, such as distance to urban points of interest. Along with the real estate listings, the data product also includes neighborhood boundaries for each city. The data product is offered as a fully documented R package and is available for scientific and educational purposes, particularly for geo-spatial studies"

keywords: Housing market; idealista.com; geo-referenced data; point-level data; open data; Spain
classoption:
  - Royal
  - times
bibliography: bibEPB.bib
bibliographystyle: sageh

@@

The data set is distributed as an R package, name {idealista18}, which can be accessed from a public Github repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}. The database idealista18 is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}. For transparency, we also share the randomisation process applied to the original data in the aforementioned Github repository.
